Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-14 write-ups/PLDI/paper.tex
Edit overhead

Getting this non-functional requirement right depends on the
constraints of the tasks to be co-scheduled with a cooperative kernel,
on properties of the cooperative kernel itself, and on the efficiency
of the GPU that is used for execution.  In \mysec\ref{sec:representativeworkloads} we conduct
experiments to understand the response rate that would be required to
co-schedule graphics rendering with a cooperative kernel in order to
achieve a smooth frame rate, and we evaluate responsiveness in
practice in \mysec\ref{sec:responsiveness}.

@@

\subsection{Sizing Non-Persistent Workloads}\label{sec:representativeworkloads}

\HEComment{given the paper title, maybe ``non-blocking'' would be more
  appropriate than ``non-persistent''? Or ``non-cooperative'' ?}

Cooperative kernels enables other \emph{non-persistent} workloads to
access GPU resources without stopping the blocking algorithm
computation entirely. We size these workloads to represent
approximatively the workload required by a system to maintain smooth
screen animations.

\paragraph{Experimental setup} \ADComment{Describe the data we are
  gathering.}  We repetitively run a kernel with a controlled workload
and detect from which workload size the screen animation starts to
jitter. Given the kernel workload and its scheduling frequency, we
deduce what is the workload and frequency needed by the system to
maintain a smooth animation. We use this approach to measure required
by the system to maintain smooth animations of three types:
(\emph{light}) emphasis of desktop icons under the mouse pointer,
(\emph{medium}) window dragging over the desktop, and (\emph{heavy})
WebGL animations in a browser.

\begin{table}
\small
\centering
\begin{tabular}{ l c c }
Workload & Duration (ms) & Period (ms)\\
\hline
light & 3 & 70\\
medium & 3 & 40\\
heavy & 10 & 40\\
\end{tabular}
\caption{Caracteristics of non-persitent workloads}
\label{tab:nonpersistentworkload}
\end{table}

\paragraph{Results}
\mytab~\ref{tab:nonpersistentworkload} illustrates the caracteristics
found for the three types of workloads. For the medium and heavy
workloads, the 40ms period corresponds to an execution every $25^{th}$
of seconds which coincide with the human persistence of vision. The
3ms execution duration of both light and medium configurations may
indicate a lower bound of GPU execution time for system workloads.

@@

\paragraph{Experimental setup}
Cooperative scheduling primitives entails an overhead even if no
killing, forking or resizing ever happen. We measure this overhead via
the execution time ratio between the original kernel and the
cooperative one, forcing the scheduler to never modify the number of
active workgroups. Since the cooperative kernel contains the
non-persistent task code anyway, its occupancy bound may be less than
the original kernel; we make sure to use this occupancy bound as the
number of workgroups for the original kernel in order to compare
relevant execution times. \HEComment{Do we need to mention the
  possible occupancy reduction as an overhead?}

\paragraph{Results}
\mytab~\ref{tab:overhead} lists overhead slowdown ratios, both
geometric mean and maximum value. The overhead is below a 2 times
slowdown even in the worst case. We consider these results as
acceptable, given that our SVM-based scheduler performance is likely a
far upper bound of what a scheduler implemented directly in the
runtime could offer.

>>>>

Getting this non-functional requirement right depends on the
constraints of the tasks to be co-scheduled with a cooperative kernel,
on properties of the cooperative kernel itself, and on the efficiency
of the GPU that is used for execution.  In \mysec\ref{sec:sizingnoncoop} we conduct
experiments to understand the response rate that would be required to
co-schedule graphics rendering with a cooperative kernel in order to
achieve a smooth frame rate, and we evaluate responsiveness in
practice in \mysec\ref{sec:responsiveness}.

@@

\subsection{Sizing Non-Cooperative Kernels}\label{sec:sizingnoncoop}

A cooperative kernel can yield workgroups to let a non-cooperative,
transient one execute. We size non-cooperative kernels to represent
approximatively the workload required by a system to maintain smooth
screen animations.

\paragraph{Experimental setup} We repetitively launch a kernel with a
controlled workload and detect from which workload size the screen
animation starts to jitter. Given the kernel workload and its
scheduling frequency, we can deduce what is the workload and frequency
needed by the system to maintain a smooth animation. We use this
approach to measure the system workloads needed to maintain smooth
animations of three types: (\emph{light}) emphasis of desktop icons
under the mouse pointer, (\emph{medium}) window dragging over the
desktop, and (\emph{heavy}) WebGL animations in a browser.

\begin{table}
\small
\centering
\begin{tabular}{ l c c }
Workload & Duration (ms) & Period (ms)\\
\hline
light & 3 & 70\\
medium & 3 & 40\\
heavy & 10 & 40\\
\end{tabular}
\caption{Caracteristics of non-cooperative workloads}
\label{tab:noncooperativeworkload}
\end{table}

\paragraph{Results}
\mytab~\ref{tab:noncooperativeworkload} illustrates the caracteristics
found for the three types of workloads. For the medium and heavy
workloads, the 40ms period corresponds to an execution every $25^{th}$
of seconds, which coincides with the human persistence of vision. The
3ms execution duration of both light and medium configurations may
indicate a lower bound of GPU execution time for system workloads.

@@

\paragraph{Experimental setup}
Cooperative scheduling primitives entails an overhead even if no
killing, forking or resizing never happens. We measure this overhead
via the execution time slowdown between the original kernel and the
cooperative one, forcing the scheduler to never modify the number of
active workgroups. Since the cooperative kernel contains the
non-cooperative kernel code anyway, its occupancy bound may be less
than the original kernel; we make sure to use this occupancy bound as
the number of workgroups for the original kernel in order to compare
relevant execution times. \HEComment{Do we need to mention the
  possible occupancy reduction as an overhead?}

\paragraph{Results}
\mytab~\ref{tab:overhead} lists both geometric mean and maximum value
of measured slowdowns. The overhead is below a 2 times slowdown even
in the worst case, and quite light in general. We consider these
results as encouraging, given that our SVM-based scheduler performance
is likely a far upper bound of what a scheduler implemented directly
in the runtime could offer.
